---
title: "Assignment 3"
author: Cynthia Begay 
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup}
knitr::opts_chunk$set(include  = TRUE)
```

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(tidytext)
library(tidyverse)
library(stringr)
library(httr)
```

##APIs

##Question 1: Using the NCBI API, look for papers that show up under the term "sars-cov-2 trial vaccine." 
Look for the data in the pubmed database, and then retrieve the details of the paper as shown in lab 7. How many papers were you able to find?

```{r}
website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2+trial+vaccine")

# Finding the counts
counts <- xml2::xml_find_first(website, "/html/body/main/div[9]/div[2]/div[2]/div[1]/span")

# Turning it into text
counts <- as.character(counts)

# Extracting the data using regex
stringr::str_extract(counts, "[0-9,]+")

```
The pubmed database retrieved 560 papers that included the terms sars, cov-2 trial vaccine. 

##Question 2:Using the list of pubmed ids you retrieved, download each papers’ details using the query parameter rettype = abstract. If you get more than 250 ids, just keep the first 250. 

```{r papers-covid-vaccine, eval=TRUE}
query_ids <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(
          db= "pubmed",
          term= "sars-cov-2 trial vaccine",
          retmax= 1000)
)
query_ids
# Extracting the content of the response of GET
ids <- httr::content(query_ids)
ids
```

```{r get-ids, eval = TRUE}
# Turn the result into a character vector
ids <- as.character(ids)
# Find all the ids 
ids <- stringr::str_extract_all(ids, "<Id>[0-9]+</Id>")[[1]]

# Remove all the leading and trailing <Id> </Id>. Make use of "|"
ids <- stringr::str_remove_all(ids, "<Id>|</Id>")
#ids <- stringr::str_remove_all(ids, "</?Id>")
```

```{r get-abstracts, eval = TRUE}
publications <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi",
  query = list(
    db = "pubmed",
    id = paste(ids, collapse = ","),
    retmax= 1000,
    rettype= "abstract"
    )
)
# Turning the output into character vector
publications <- httr::content(publications)
publications_txt <- as.character(publications)
```
##Question 3: Create a dataset containing the following:

Pubmed ID number,
Title of the paper,
Name of the journal where it was published,
Publication date, and
Abstract of the paper (if any).

```{r}
pub_char_list <- xml2::xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)
```

```{r process-abstracts, eval = TRUE}
abstracts <- str_extract(pub_char_list, "<Abstract>(\\n|.)+</Abstract>")
abstracts <- str_remove_all(abstracts, "</?[[:alnum:]]+>")
abstracts <- str_replace_all(abstracts, "\\s+", " ")
table(is.na(abstracts))
```

```{r process-titles, eval = TRUE}
titles <- str_extract(pub_char_list, "<ArticleTitle>(\\n|.)+</ArticleTitle>")
titles <- str_remove_all(titles, "</?[[:alnum:]]+>")
titles <- str_replace_all(titles, "\\s+", " ")
table(is.na(titles))
```

```{r process-journal, eval = TRUE}
journal <- str_extract(pub_char_list, "<Title>(\\n|.)+</Title>")
journal <- str_remove_all(journal, "</?[[:alnum:]]+>")
journal <- str_replace_all(journal, "\\s+", " ")
table(is.na(journal))
```

```{r process-pubdate, eval = TRUE}
pubdate <- str_extract(pub_char_list, "<PubDate>(\\n|.)+</PubDate>")
pubdate <- str_remove_all(pubdate, "</?[[:alnum:]]+>")
pubdate <- str_replace_all(pubdate, "\\s+", " ")
table(is.na(pubdate))
```

```{r build-db, eval = TRUE}
database <- data.frame(
  PubMedID = ids,
  Title = titles,
  Journal = journal,
  Pubdate = pubdate,
  Abstracts = abstracts
)
knitr::kable(database)
```

## Text Mining

```{r, warning=FALSE, message=FALSE}
pubmed_samples <- read_csv("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/03_pubmed/pubmed.csv")

head(pubmed_samples)

pubmed_samples$abstract[1]
```

---

## Question 1: Tokenize the abstracts and count the number of each token.
Do you see anything interesting? Does removing stop words change what tokens appear as the most frequent? What are the 5 most common tokens for each search term after removing stopwords?

```{r}
#Tokenize abstracts and count the number of each token
pubmed_samples %>%
  unnest_tokens(output = word, input = abstract) %>%
  count(word, term, sort=TRUE) %>%
  top_n(n = 20, wt = n) %>%
  ggplot(aes(x = n, y = fct_reorder(word, n))) +
  geom_col()
```
There are 20,567 tokens. Viewing the top 20 most common tokens, most are stop words. The first nouns are Covid, patients, cancer, prostate.
Top 5 words: the, of, and, in, to
```{r}
stop_words
number_words <-as.character(seq(0, 100))
```

```{r}
#Remove stopwords 
pubmed_samples %>%
 unnest_tokens(word, abstract) %>%
  filter(!(word%in% as.character(seq(0,100))))%>%
  anti_join(stop_words, by = c("word")) %>%
  count(word, sort = TRUE) %>%
  top_n(10, n) %>%
  ggplot(aes(n, fct_reorder(word, n))) +
  geom_col()
```
Removing stop words shows a more descriptive result. The top 5 hits are covid, patients, cancer, prostate, and disease. 
```{r}
pubmed_Q1 <- pubmed_samples %>% 
  unnest_tokens(word, abstract) %>%
  anti_join(tidytext :: stop_words, by = c("word")) %>%
  group_by(term) %>%
  count(word) %>%
  top_n(5,n)
```

```{r}
pubmed_Q1
```

Covid: 19, covid, disease, pandemic, patients
cystic fibrosis: cf, cystic, disease, fibrosis, patients
meningitis: clinical, csf, 	meningeal, meningitis, patients
preeclampsia: eclampsia, pre, preeclampsia, pregnancy, women
Prostate Cancer: cancer, disease, patients, prostate, treatment
---

## Question 2: Tokenize the abstracts into bigrams. 
Find the 10 most common bigram and visualize them with ggplot2.

```{r}
pubmed_samples %>%
  unnest_ngrams(output = token, input = abstract, n = 2) %>%
  count(token, sort=TRUE) %>%
  top_n(n = 10, wt = n) %>%
  ggplot(aes(x = n, y = fct_reorder(token, n))) +
  geom_col()
```

---

## Question 3: Calculate the TF-IDF value for each word-search term combination.
Calculate the TF-IDF value for each word-search term combination. (here you want the search term to be the “document”) What are the 5 tokens from each search term with the highest TF-IDF value? How are the results different from the answers you got in question 1?

```{r}
pubmed_samples %>%
 unnest_tokens(word, abstract) %>%
  count(word, term) %>%
  bind_tf_idf(word, term, n) %>%
  arrange(desc(tf_idf))

```

```{r}
pubmed_Q3 <- pubmed_samples %>% 
  unnest_tokens(word, abstract) %>%
  group_by(term) %>%
  count(word, term) %>%
  bind_tf_idf(word, term, n) %>%
  top_n(5,tf_idf) %>%
  arrange(desc(tf_idf))
```
```{r}
pubmed_Q3
```

The words with the highest tf_idf values are more descriptive/unique than the words in question 1. For example, covid's words with the highest tf_idf values are: covid, pandemic, coronavirus, sars, and cov. 

covid: covid, pandemic, coronavirus, sars, cov
cystic fibrosis: cf, fibrosis, cystic, cftr, sweat
meningitis: meningitis, meningeal, pachymeningitis, csf, meninges
preeclampsia: eclampsia, preeclampsia, pregnancy, maternal, gestational
prostate cancer: prostate cancer, androgen, psa, prostatectomy, castration


