---
title: "Assignment 4"
author: Cynthia Begay 
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, eval=TRUE}
a = matrix(1, 4, 5)

# Total row sums
fun1 <- function(mat) {
  n <- nrow(mat)
  ans <- double(n)
  for (i in 1:n) {
    ans[i] <- sum(mat[i, ])
  }
  ans
}

fun1alt <- function(mat) {
  x = rowSums(mat, na.rm = TRUE)
  #x = apply(mat, 1, sum)
  x
}
```

```{r problem 1, eval=FALSE}
fun1alt(dat)
fun1(dat)
```

```{r}
# Cumulative product by row
fun2 <- function(mat) {
  n <- nrow(mat)
  k <- ncol(mat)
  ans <- mat
  for (i in 1:n) {
    for (j in 2:k)
      ans[i,j] <- mat[i, j] + ans[i, j - 1]
    }
  ans
 }


fun2alt <- function(mat) {
  x = apply(mat, 1, FUN = cumsum)
  t(x)
}
```

```{r}
# Use the data with this code
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)

# Test for the first
microbenchmark::microbenchmark(
  fun1(dat),
  fun1alt(dat), unit = "relative", check = "equivalent"
)

# Test for the second
microbenchmark::microbenchmark(
  fun2(dat),
  fun2alt(dat), unit = "relative", check = "equivalent"
)
```


```{r problem 2, eval = TRUE}
sim_pi <- function(n = 1000, i = NULL) {
  p <- matrix(runif(n*2), ncol = 2)
  mean(rowSums(p^2) < 1) * 4
}

# Here is an example of the run
set.seed(156)
sim_pi(1000) # 3.132
```

```{r}
# This runs the simulation a 4,000 times, each with 10,000 points
set.seed(1231)
system.time({
  ans <- unlist(lapply(1:4000, sim_pi, n = 10000))
  print(mean(ans))
})
```

```{r}

library(parallel)
cl <- makePSOCKcluster(4L)
clusterSetRNGStream(cl, 1231) #Equivalent to set.seed(1231)

system.time({
  ans <- unlist(parLapply(
    cl = cl,
    1:4000,
    sim_pi,
    n=10000
  ))
  
  parallel::stopCluster(cl)
  
  print(mean(ans))
})
```

```{r}
#### Part 2: SQL #####
#install.packages(c("RSQLite", "DBI"))
library(RSQLite)
library(DBI)

# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
film <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film.csv")
film_category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film_category.csv")
category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/category.csv")

# Copy data.frames to database
dbWriteTable(con, "film", film)
dbWriteTable(con, "film_category", film_category)
dbWriteTable(con, "category", category)
```

```{sql, connection=con}
PRAGMA table_info(film)
```
```{sql, connection=con}
PRAGMA table_info(film_category)
```

```{sql, connection=con}
PRAGMA table_info(category)
```
Question 1: How many movies are there available in each rating category? 
```{sql, connection=con}
SELECT film_id, title, rating
FROM film
ORDER by rating
```
There are 1000 observations with an associated rating.

```{sql, connection=con}
SELECT film_id, title, rating
FROM film
WHERE rating IN ('G')
```
There are 180 movies with a G rating. 

```{sql, connection=con}
SELECT film_id, title, rating
FROM film
WHERE rating IN ('PG')
```
There are 194 movies with a PG rating. 

```{sql, connection=con}
SELECT film_id, title, rating
FROM film
WHERE rating IN ('PG-13')
```
There are 223 movies with a PG-13 rating. 

```{sql, connection=con}
SELECT film_id, title, rating
FROM film
WHERE rating IN ('R')
```
There are 195 movies with an R rating. 

```{sql, connection=con}
SELECT film_id, title, rating
FROM film
WHERE rating IN ('NC-17')
```
There are 210 movies with an NC-17 rating.

Question 2: What is the average replacement cost and rental rate for each rating category.

```{sql, connection=con}
SELECT rating,
       AVG(rental_rate) AS avg_rate,
       AVG(replacement_cost) AS avg_replace
FROM film
GROUP BY rating
```

The average rental rates and average replacement costs by ratings are: 
Rated G, average rental rate = $2.91, average replacement cost = $20.12
Rated PG, average rental rate = $3.05, average replacement cost = $18.96
Rated PG-13, average rental rate = $3.03, average replacement cost = $20.40
Rated R, average rental rate = $2.93, average replacement cost = $20.23
Rated NC-17, average rental rate = $2.97, average replacement cost = $20.14


Question 3: Use table film_category together with film to find the how many films there are within each category ID

```{sql, connection=con}
SELECT film.film_id, film_category.category_id
  FROM film
  INNER JOIN film_category ON film.film_id = film_category.film_id
  GROUP BY category_id;
```
There are 16 categories total. The number of films per category are: 
Category 1: 19
Category 2: 18
Category 3: 48
Category 4: 14
Category 5: 7
Category 6: 1
Category 7: 33
Category 8: 5
Category 9: 6
Category 10: 46
Category 11: 2
Category 12: 12
Category 13: 22
Category 14: 26
Category 15: 10
Category 16: 41

Question 4: 
Incorporate table category into the answer to the previous question to find the name of the most popular category.
```{sql, connection=con}
SELECT film.film_id, film_category.category_id, category.name
  FROM ((film
  INNER JOIN film_category ON film.film_id = film_category.film_id)
  INNER JOIN category ON film_category.category_id = category.category_id)
  GROUP BY name;
```
The most popular category is catgory 3 "Children" with 48 movies. 







